---
title: "stats100_project_update"
output:
  word_document: default
  pdf_document: default
date: "2024-02-24"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(dplyr)
library(purrr)
clutch_data_20_21 <- read.csv("player_clutch_2020-2021.csv")
regular_season_data_20_21 <- read.csv("player_regular_season_2020-2021.csv")
# TODO: add data from other seasons and re-run analysis
clutch_data_18_19 <- read.csv("player_clutch_2018-2019.csv") 
regular_season_data_18_19 <- read.csv("player_regular_season_2018-2019.csv")
clutch_data_19_20 <- read.csv("player_clutch_2019-2020.csv")
regular_season_data_19_20 <- read.csv("player_regular_season_2019-2020.csv")
clutch_data_21_22 <- read.csv("player_clutch_2021-2022.csv")
regular_season_data_21_22 <- read.csv("player_regular_season_2021-2022.csv")
clutch_data_22_23 <- read.csv("player_clutch_2022-2023.csv")
regular_season_data_22_23 <- read.csv("player_regular_season_2022-2023.csv")
```

```{r}
clutch_data_18_19 <- clutch_data_18_19 %>% select(-(contains('RANK') | contains('FANTASY')))
clutch_data_19_20 <- clutch_data_19_20 %>% select(-(contains('RANK') | contains('FANTASY')))
clutch_data_20_21 <- clutch_data_20_21 %>% select(-(contains('RANK') | contains('FANTASY')))
clutch_data_21_22 <- clutch_data_21_22 %>% select(-(contains('RANK') | contains('FANTASY')))
clutch_data_22_23 <- clutch_data_22_23 %>% select(-(contains('RANK') | contains('FANTASY')))


datasets_clutch <- list(clutch_data_18_19, clutch_data_19_20 , clutch_data_20_21, clutch_data_21_22, clutch_data_22_23)
all_seasons_clutch <- reduce(datasets_clutch, ~inner_join(.x, .y, by = "PLAYER_ID"))

all_seasons_clutch <- all_seasons_clutch %>%
  mutate(
    total_W = W + W.x + W.x.x + W.y + W.y.y,
    total_L = L + L.x + L.x.x + L.y + L.y.y,
    total_GP = GP + GP.x + GP.x.x + GP.y + GP.y.y,
    AGE = (AGE + AGE.x + AGE.x.x + AGE.y + AGE.y.y) / 5
  )

weighted_avg <- function(seasons_data, stat_name) {
  weights <- seasons_data %>% select(GP, GP.x, GP.x.x, GP.y, GP.y.y) %>% as.matrix()
  pattern <- paste0("^", stat_name, "(\\.x|\\.x\\.x|\\.y|\\.y\\.y)?$")
  stats <- seasons_data %>% select(matches(pattern)) %>% as.matrix()
  weighted_mean <- rowSums(weights * stats, na.rm = TRUE) / rowSums(weights, na.rm = TRUE)
  return(weighted_mean)
}

all_seasons_clutch$W_PCT <- weighted_avg(all_seasons_clutch, "W_PCT")
all_seasons_clutch$W_MIN <- weighted_avg(all_seasons_clutch, "MIN")
all_seasons_clutch$FGM <- weighted_avg(all_seasons_clutch, "FGM")
all_seasons_clutch$FGA <- weighted_avg(all_seasons_clutch, "FGA")
all_seasons_clutch$FG_PCT <- weighted_avg(all_seasons_clutch, "FG_PCT")
all_seasons_clutch$FG3M <- weighted_avg(all_seasons_clutch, "FG3M")
all_seasons_clutch$FG3A <- weighted_avg(all_seasons_clutch, "FG3A")
all_seasons_clutch$FG3_PCT <- weighted_avg(all_seasons_clutch, "FG3_PCT")
all_seasons_clutch$FTM <- weighted_avg(all_seasons_clutch, "FTM")
all_seasons_clutch$FTA <- weighted_avg(all_seasons_clutch, "FTA")
all_seasons_clutch$FT_PCT <- weighted_avg(all_seasons_clutch, "FT_PCT")
all_seasons_clutch$OREB <- weighted_avg(all_seasons_clutch, "OREB")
all_seasons_clutch$DREB <- weighted_avg(all_seasons_clutch, "DREB")
all_seasons_clutch$REB <- weighted_avg(all_seasons_clutch, "REB")
all_seasons_clutch$AST <- weighted_avg(all_seasons_clutch, "AST")
all_seasons_clutch$TOV <- weighted_avg(all_seasons_clutch, "TOV")
all_seasons_clutch$STL <- weighted_avg(all_seasons_clutch, "STL")
all_seasons_clutch$BLK <- weighted_avg(all_seasons_clutch, "BLK")
all_seasons_clutch$PF <- weighted_avg(all_seasons_clutch, "PF")
all_seasons_clutch$PFD <- weighted_avg(all_seasons_clutch, "PFD")
all_seasons_clutch$PTS <- weighted_avg(all_seasons_clutch, "PTS")
all_seasons_clutch$PLUS_MINUS <- weighted_avg(all_seasons_clutch, "PLUS_MINUS")

all_seasons_clutch <- all_seasons_clutch %>%
  select('PLAYER_ID', 'PLAYER_NAME', 'TEAM_ID', 'TEAM_ABBREVIATION', 'AGE', 'total_GP', 'total_W', 'total_L', "W_PCT", "MIN", "FGM", "FGA", "FG_PCT", "FG3M", "FG3A", "FG3_PCT", "FTM", "FTA", "FT_PCT", "OREB", "DREB", "REB", "AST", "TOV", "STL", "BLK", "PF", "PFD", "PTS", "PLUS_MINUS")

```

```{r}
regular_season_data_18_19 <- regular_season_data_18_19 %>% select(-(contains('RANK') | contains('FANTASY')))
regular_season_data_19_20 <- regular_season_data_19_20 %>% select(-(contains('RANK') | contains('FANTASY')))
regular_season_data_20_21 <- regular_season_data_20_21 %>% select(-(contains('RANK') | contains('FANTASY')))
regular_season_data_21_22 <- regular_season_data_21_22 %>% select(-(contains('RANK') | contains('FANTASY')))
regular_season_data_22_23 <- regular_season_data_22_23 %>% select(-(contains('RANK') | contains('FANTASY')))


datasets_regular <- list(regular_season_data_18_19, regular_season_data_19_20 , regular_season_data_20_21, regular_season_data_21_22, regular_season_data_22_23)
all_regular_seasons <- reduce(datasets_regular, ~inner_join(.x, .y, by = "PLAYER_ID"))

all_regular_seasons <- all_regular_seasons %>%
  mutate(
    total_W = W + W.x + W.x.x + W.y + W.y.y,
    total_L = L + L.x + L.x.x + L.y + L.y.y,
    total_GP = GP + GP.x + GP.x.x + GP.y + GP.y.y,
    AGE = (AGE + AGE.x + AGE.x.x + AGE.y + AGE.y.y) / 5
  )

weighted_avg <- function(seasons_data, stat_name) {
  weights <- seasons_data %>% select(GP, GP.x, GP.x.x, GP.y, GP.y.y) %>% as.matrix()
  pattern <- paste0("^", stat_name, "(\\.x|\\.x\\.x|\\.y|\\.y\\.y)?$")
  stats <- seasons_data %>% select(matches(pattern)) %>% as.matrix()
  weighted_mean <- rowSums(weights * stats, na.rm = TRUE) / rowSums(weights, na.rm = TRUE)
  return(weighted_mean)
}

all_regular_seasons$W_PCT <- weighted_avg(all_regular_seasons, "W_PCT")
all_regular_seasons$MIN <- weighted_avg(all_regular_seasons, "MIN")
all_regular_seasons$FGM <- weighted_avg(all_regular_seasons, "FGM")
all_regular_seasons$FGA <- weighted_avg(all_regular_seasons, "FGA")
all_regular_seasons$FG_PCT <- weighted_avg(all_regular_seasons, "FG_PCT")
all_regular_seasons$FG3M <- weighted_avg(all_regular_seasons, "FG3M")
all_regular_seasons$FG3A <- weighted_avg(all_regular_seasons, "FG3A")
all_regular_seasons$FG3_PCT <- weighted_avg(all_regular_seasons, "FG3_PCT")
all_regular_seasons$FTM <- weighted_avg(all_regular_seasons, "FTM")
all_regular_seasons$FTA <- weighted_avg(all_regular_seasons, "FTA")
all_regular_seasons$FT_PCT <- weighted_avg(all_regular_seasons, "FT_PCT")
all_regular_seasons$OREB <- weighted_avg(all_regular_seasons, "OREB")
all_regular_seasons$DREB <- weighted_avg(all_regular_seasons, "DREB")
all_regular_seasons$REB <- weighted_avg(all_regular_seasons, "REB")
all_regular_seasons$AST <- weighted_avg(all_regular_seasons, "AST")
all_regular_seasons$TOV <- weighted_avg(all_regular_seasons, "TOV")
all_regular_seasons$STL <- weighted_avg(all_regular_seasons, "STL")
all_regular_seasons$BLK <- weighted_avg(all_regular_seasons, "BLK")
all_regular_seasons$PF <- weighted_avg(all_regular_seasons, "PF")
all_regular_seasons$PFD <- weighted_avg(all_regular_seasons, "PFD")
all_regular_seasons$PTS <- weighted_avg(all_regular_seasons, "PTS")
all_regular_seasons$PLUS_MINUS <- weighted_avg(all_regular_seasons, "PLUS_MINUS")

all_regular_seasons <- all_regular_seasons %>%
  select('PLAYER_ID', 'PLAYER_NAME', 'TEAM_ID', 'TEAM_ABBREVIATION', 'AGE', 'total_GP', 'total_W', 'total_L', "W_PCT", "MIN", "FGM", "FGA", "FG_PCT", "FG3M", "FG3A", "FG3_PCT", "FTM", "FTA", "FT_PCT", "OREB", "DREB", "REB", "AST", "TOV", "STL", "BLK", "PF", "PFD", "PTS", "PLUS_MINUS")

```


```{r}
all_seasons_clutch <- mutate(all_seasons_clutch, clutch_PER = (85.910 * FGM + 53.897 * STL + 51.757 * FG3M + 46.845 * FTM + 39.190 * BLK + 39.190 * OREB + 34.677 * AST + 14.707 * DREB - 17.174 * PF - 20.091 * (FTA-FTM) - 39.190 * (FGA -FGM) - 53.897 * TOV)/MIN)

all_regular_seasons <- mutate(all_regular_seasons, PER = (85.910 * FGM + 53.897 * STL + 51.757 * FG3M + 46.845 * FTM + 39.190 * BLK + 39.190 * OREB + 34.677 * AST + 14.707 * DREB - 17.174 * PF - 20.091 * (FTA-FTM) - 39.190 * (FGA -FGM) - 53.897 * TOV)/MIN)

full_df <- inner_join(all_seasons_clutch, all_regular_seasons, by = "PLAYER_ID")

updated_df <- full_df[full_df$total_GP.x * full_df$MIN.x >= 180, ]
```

```{r}
library(ggplot2)

model <- lm(clutch_PER ~ PER, data = updated_df)

summary_model <- summary(model)

r_squared <- summary_model$r.squared
print(r_squared)

ggplot(updated_df, aes(x = PER, y = clutch_PER)) +
  geom_point() +  # Add points for the scatter plot
  geom_smooth(method = "lm", se = FALSE, color = "red") +  # Add a line of best fit
  labs(x = "PER", y = "Clutch PER", title = "Regular Season PER vs Clutch Time PER") + 
  theme(plot.title = element_text(hjust = 0.5)) 
```

```{r}
updated_df <- mutate(updated_df, clutch_elevation = clutch_PER - PER)
updated_df <- mutate(updated_df, clutch_elevator = ifelse(clutch_elevation > 0, 1, 0))
```

Visualizing clutch elevation:
```{r}

mean_value = mean(updated_df$clutch_elevation)
sd_value = sd(updated_df$clutch_elevation)

ggplot(updated_df, aes(x = clutch_elevation)) + 
  geom_histogram(binwidth = 1, fill = "blue", color = "black") + # Adjust binwidth as needed
  ggtitle("Histogram of Clutch Elevation") + 
  xlab("Clutch Elevation") + 
  ylab("Frequency") +
  theme(plot.title = element_text(hjust = 0.5)) +
  annotate("text", x = Inf, y = Inf, label = paste("Mean =", round(mean_value, 2)), hjust = 1.1,vjust = 3, color="black", size = 4.5) +
  annotate("text", x = Inf, y = Inf, label = paste("SD =", round(sd_value, 2)), hjust = 1.1,vjust = 1.5, color="black", size = 4.5) +
  theme(plot.margin = margin(5.5, 40, 5.5, 5.5))

print(median(updated_df$clutch_elevation))

highest_clutch_elevation_players <- updated_df %>%
  arrange(desc(clutch_elevation)) %>%
  head(6)

lowest_clutch_elevation_players <- updated_df %>%
  arrange(clutch_elevation) %>%
  head(6)

highest_clutch_PER_players <- updated_df %>%
  arrange(desc(clutch_PER)) %>%
  head(10)

highest_clutch_players <- updated_df %>%
  arrange(desc(clutch_PER + clutch_elevation)) %>%
  head(10)

# Combine these two subsets
selected_players <- bind_rows(lowest_clutch_elevation_players, highest_clutch_elevation_players)

# Add a new column to indicate if they are in the lowest or highest group for coloring
selected_players$group <- rep(c("Lowest", "Highest"), each = 6)

# Plotting
ggplot(selected_players, aes(x = reorder(PLAYER_NAME.x, clutch_elevation), y = clutch_elevation, fill = group)) +
  geom_bar(stat = "identity", show.legend = FALSE) +
  scale_fill_manual(values = c("Lowest" = "red", "Highest" = "green")) +
  coord_flip() + # Flip coordinates to make the names vertical (alternative to making them diagonal)
  theme(axis.text.y = element_text(angle = 45, hjust = 1)) + # Adjust text angle and alignment
  labs(title = "Players with the Highest and Lowest Clutch Elevation", x = "Player Name", y = "Clutch Elevation") + 
  theme(plot.title = element_text(hjust = 0.5)) 
```

## Finding: on average, players tends to perform worse in the clutch. Normally distributed centered around a slightly negative mean. 

```{r}
# Fit multiple linear regression model
full_model <- lm(clutch_elevation ~ AGE.y + FG_PCT.y + FG3_PCT.y + FT_PCT.y + TOV.y + PF.y + OREB.y + W_PCT.y + MIN.y + FGA.y + FG3A.y + FTA.y + DREB.y + AST.y + STL.y + BLK.y + PLUS_MINUS.y, data = updated_df)

# Summarize the model
summary_model <- summary(full_model)

r_squared <- summary_model$r.squared
adj_r_squared <- summary_model$adj.r.squared
print(r_squared)
print(summary_model)

updated_df$predicted_clutch_elevation <- predict(full_model)

# Create scatter plot matrix with fitted values
ggplot(updated_df, aes(x = predicted_clutch_elevation, y = clutch_elevation)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE, color = "red")  + 
  labs(x = "Predicted Clutch Elevation", y = "Observed Clutch Elevation", title = "Predicted vs Observed Clutch Elevation") +
  theme(plot.title = element_text(hjust = 0.5)) +
  annotate("text", x = Inf, y = Inf, label = paste("Multiple R-squared =", round(r_squared, 3)), 
           hjust = 1.1, vjust = 2, color = "black") +
  annotate("text", x = Inf, y = Inf, label = paste("Adjusted R-squared =", round(adj_r_squared, 3)), 
           hjust = 1.1, vjust = 3.5, color = "black") +
  theme(plot.margin = margin(5.5, 40, 5.5, 5.5)) 
```
Performing feature selection with stepwise regression
```{r}
library(MASS)

stepModel <- stepAIC(full_model, direction = "both", trace = FALSE)
summary(stepModel)
```

```{r}
library(caTools)
library(caret)
library(Metrics)
library(MASS)

set.seed(1) 
splitIndex <- createDataPartition(updated_df$clutch_elevation, p = .75, list = FALSE)
train_data <- updated_df[splitIndex,]
test_data <- updated_df[-splitIndex,]


# Fit multiple linear regression model
full_model <- lm(clutch_elevation ~ AGE.y + FG_PCT.y + FG3_PCT.y + FT_PCT.y + TOV.y + PF.y + OREB.y + W_PCT.y + MIN.y + FGA.y + FG3A.y + FTA.y + DREB.y + AST.y + STL.y + BLK.y + PLUS_MINUS.y, data = train_data)

stepModel <- stepAIC(full_model, direction = "both", trace = FALSE)

predictions <- predict(stepModel, newdata=test_data)

rmse_value <- rmse(test_data$clutch_elevation, predictions)

r_squared <- summary(stepModel)$r.squared
adj_r_squared <- summary_model$adj.r.squared
print(summary(stepModel))

test_data$predicted_clutch_elevation <- predictions

# Create scatter plot matrix with fitted values
ggplot(test_data, aes(x = predicted_clutch_elevation, y = clutch_elevation)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE, color = "red")  + 
  labs(x = "Predicted Clutch Elevation", y = "Observed Clutch Elevation", title = "Predicted vs. Observed Clutch Elevation on Test Set") +
  theme(plot.title = element_text(hjust = 0.5)) +
  annotate("text", x = Inf, y = Inf, label = paste("Multiple R-squared =", round(r_squared, 3)), 
           hjust = 1.1, vjust = 5.1, color = "black") +
  annotate("text", x = Inf, y = Inf, label = paste("Adjusted R-squared =", round(adj_r_squared, 3)), 
           hjust = 1.1, vjust = 6.6, color = "black") +
  theme(plot.margin = margin(5.5, 40, 5.5, 5.5)) 
```

# Major Finding: free throw percentage is the statistic that is consistently most signficiantly positively correlated with clutch elevation. Though overall predicitve power is still very low, seems that free throw pct is the closest proxy. 


```{r}
model_ft <- lm(clutch_elevation ~ FT_PCT.y, data = updated_df)

summary_model <- summary(model_ft)

r_squared <- summary_model$r.squared

cor_coef = cor(updated_df$FT_PCT.y, updated_df$clutch_elevation)

ggplot(updated_df, aes(x = FT_PCT.y, y = clutch_elevation)) +
  geom_point() +  
  geom_smooth(method = "lm", se = FALSE, color = "red") +  # Add a line of best fit
  labs(x = "Regular Season Free Throw Percent", y = "Clutch Elevation", title = "Free Throw Percent vs. Clutch Elevation") + 
  theme(plot.title = element_text(hjust = 0.5)) +
  annotate("text", x = Inf, y = Inf, label = paste("Correlation =", round(cor_coef, 3)), 
           hjust = 4.35, vjust = 2, color = "black", size = 4.5)

```



# Maybe something with logistic classification for clutch riser/faller/stays the same
```{r}
# Splitting data
set.seed(1) # For reproducibility
splitIndex <- createDataPartition(updated_df$clutch_elevator, p = .75, list = FALSE)
trainData <- updated_df[splitIndex,]
testData <- updated_df[-splitIndex,]

classifier_model <- glm(clutch_elevator ~ AGE.y + FG_PCT.y + FG3_PCT.y + FT_PCT.y + TOV.y + PF.y + OREB.y + W_PCT.y + MIN.y + FGA.y + FG3A.y + FTA.y + DREB.y + AST.y + STL.y + BLK.y + PLUS_MINUS.y, data = trainData, family = binomial)

predictions <- predict(classifier_model, newdata = testData, type = "response")
predictedClass <- ifelse(predictions > 0.5, 1, 0)

correct_predictions <- predictedClass == testData$clutch_elevator
accuracy <- sum(correct_predictions) / length(correct_predictions)
print(accuracy)
print(summary(classifier_model))

classifier_stepModel <- stepAIC(classifier_model, direction = "both", trace = FALSE)

print(summary(classifier_stepModel))

predictions <- predict(classifier_stepModel, newdata = testData, type = "response")
predictedClass <- ifelse(predictions > 0.5, 1, 0)
correct_predictions <- predictedClass == testData$clutch_elevator
accuracy <- sum(correct_predictions) / length(correct_predictions)
print(accuracy)
```






```{r}
playoffs_clutch_data_18_19 <- read.csv("playoffs_clutch_2018-2019.csv") 
playoffs_data_18_19 <- read.csv("playoffs_2018-2019.csv")
playoffs_clutch_data_19_20 <- read.csv("playoffs_clutch_2019-2020.csv")
playoffs_data_19_20 <- read.csv("playoffs_2019-2020.csv")
playoffs_clutch_data_20_21 <- read.csv("playoffs_clutch_2020-2021.csv")
playoffs_data_20_21 <- read.csv("playoffs_2020-2021.csv")
playoffs_clutch_data_21_22 <- read.csv("playoffs_clutch_2021-2022.csv")
playoffs_data_21_22 <- read.csv("playoffs_2021-2022.csv")
playoffs_clutch_data_22_23 <- read.csv("playoffs_clutch_2022-2023.csv")
playoffs_data_22_23 <- read.csv("playoffs_2022-2023.csv")
```

```{r}
playoffs_clutch_data_18_19 <- playoffs_clutch_data_18_19 %>% select(-(contains('RANK') | contains('FANTASY')))
playoffs_clutch_data_19_20 <- playoffs_clutch_data_19_20 %>% select(-(contains('RANK') | contains('FANTASY')))
playoffs_clutch_data_20_21 <- playoffs_clutch_data_20_21 %>% select(-(contains('RANK') | contains('FANTASY')))
playoffs_clutch_data_21_22 <- playoffs_clutch_data_21_22 %>% select(-(contains('RANK') | contains('FANTASY')))
playoffs_clutch_data_22_23 <- playoffs_clutch_data_22_23 %>% select(-(contains('RANK') | contains('FANTASY')))


playoffs_datasets_clutch <- list(playoffs_clutch_data_18_19, playoffs_clutch_data_19_20 , playoffs_clutch_data_20_21, playoffs_clutch_data_21_22, playoffs_clutch_data_22_23)
all_playoffs_clutch <- reduce(playoffs_datasets_clutch, ~inner_join(.x, .y, by = "PLAYER_ID"))

all_playoffs_clutch <- all_playoffs_clutch %>%
  mutate(
    total_W = W + W.x + W.x.x + W.y + W.y.y,
    total_L = L + L.x + L.x.x + L.y + L.y.y,
    total_GP = GP + GP.x + GP.x.x + GP.y + GP.y.y,
    AGE = (AGE + AGE.x + AGE.x.x + AGE.y + AGE.y.y) / 5
  )

weighted_avg <- function(seasons_data, stat_name) {
  weights <- seasons_data %>% select(GP, GP.x, GP.x.x, GP.y, GP.y.y) %>% as.matrix()
  pattern <- paste0("^", stat_name, "(\\.x|\\.x\\.x|\\.y|\\.y\\.y)?$")
  stats <- seasons_data %>% select(matches(pattern)) %>% as.matrix()
  weighted_mean <- rowSums(weights * stats, na.rm = TRUE) / rowSums(weights, na.rm = TRUE)
  return(weighted_mean)
}

all_playoffs_clutch$W_PCT <- weighted_avg(all_playoffs_clutch, "W_PCT")
all_playoffs_clutch$W_MIN <- weighted_avg(all_playoffs_clutch, "MIN")
all_playoffs_clutch$FGM <- weighted_avg(all_playoffs_clutch, "FGM")
all_playoffs_clutch$FGA <- weighted_avg(all_playoffs_clutch, "FGA")
all_playoffs_clutch$FG_PCT <- weighted_avg(all_playoffs_clutch, "FG_PCT")
all_playoffs_clutch$FG3M <- weighted_avg(all_playoffs_clutch, "FG3M")
all_playoffs_clutch$FG3A <- weighted_avg(all_playoffs_clutch, "FG3A")
all_playoffs_clutch$FG3_PCT <- weighted_avg(all_playoffs_clutch, "FG3_PCT")
all_playoffs_clutch$FTM <- weighted_avg(all_playoffs_clutch, "FTM")
all_playoffs_clutch$FTA <- weighted_avg(all_playoffs_clutch, "FTA")
all_playoffs_clutch$FT_PCT <- weighted_avg(all_playoffs_clutch, "FT_PCT")
all_playoffs_clutch$OREB <- weighted_avg(all_playoffs_clutch, "OREB")
all_playoffs_clutch$DREB <- weighted_avg(all_playoffs_clutch, "DREB")
all_playoffs_clutch$REB <- weighted_avg(all_playoffs_clutch, "REB")
all_playoffs_clutch$AST <- weighted_avg(all_playoffs_clutch, "AST")
all_playoffs_clutch$TOV <- weighted_avg(all_playoffs_clutch, "TOV")
all_playoffs_clutch$STL <- weighted_avg(all_playoffs_clutch, "STL")
all_playoffs_clutch$BLK <- weighted_avg(all_playoffs_clutch, "BLK")
all_playoffs_clutch$PF <- weighted_avg(all_playoffs_clutch, "PF")
all_playoffs_clutch$PFD <- weighted_avg(all_playoffs_clutch, "PFD")
all_playoffs_clutch$PTS <- weighted_avg(all_playoffs_clutch, "PTS")
all_playoffs_clutch$PLUS_MINUS <- weighted_avg(all_playoffs_clutch, "PLUS_MINUS")

all_playoffs_clutch <- all_playoffs_clutch %>%
  select('PLAYER_ID', 'PLAYER_NAME', 'TEAM_ID', 'TEAM_ABBREVIATION', 'AGE', 'total_GP', 'total_W', 'total_L', "W_PCT", "MIN", "FGM", "FGA", "FG_PCT", "FG3M", "FG3A", "FG3_PCT", "FTM", "FTA", "FT_PCT", "OREB", "DREB", "REB", "AST", "TOV", "STL", "BLK", "PF", "PFD", "PTS", "PLUS_MINUS")

```

```{r}
playoffs_data_18_19 <- playoffs_data_18_19 %>% select(-(contains('RANK') | contains('FANTASY')))
playoffs_data_19_20 <- playoffs_data_19_20 %>% select(-(contains('RANK') | contains('FANTASY')))
playoffs_data_20_21 <- playoffs_data_20_21 %>% select(-(contains('RANK') | contains('FANTASY')))
playoffs_data_21_22 <- playoffs_data_21_22 %>% select(-(contains('RANK') | contains('FANTASY')))
playoffs_data_22_23 <- playoffs_data_22_23 %>% select(-(contains('RANK') | contains('FANTASY')))


datasets_playoffs<- list(playoffs_data_18_19, playoffs_data_19_20 , playoffs_data_20_21, playoffs_data_21_22, playoffs_data_22_23)
all_playoffs <- reduce(datasets_playoffs, ~full_join(.x, .y, by = "PLAYER_ID"))

all_playoffs <- all_playoffs %>%
  mutate(
    total_W = rowSums(select(., W, W.x, W.x.x, W.y, W.y.y), na.rm = TRUE),
    total_L = rowSums(select(., L, L.x, L.x.x, L.y, L.y.y), na.rm = TRUE),
    total_GP = rowSums(select(., GP, GP.x, GP.x.x, GP.y, GP.y.y), na.rm = TRUE),
    AGE = rowMeans(select(., AGE, AGE.x, AGE.x.x, AGE.y, AGE.y.y), na.rm = TRUE),
  )
all_playoffs <- all_playoffs %>%
  mutate(PLAYER_NAME = coalesce(PLAYER_NAME, PLAYER_NAME.x, PLAYER_NAME.y, PLAYER_NAME.x.x, PLAYER_NAME.y.y))

weighted_avg <- function(seasons_data, stat_name) {
  weights <- seasons_data %>% select(GP, GP.x, GP.x.x, GP.y, GP.y.y) %>% as.matrix()
  pattern <- paste0("^", stat_name, "(\\.x|\\.x\\.x|\\.y|\\.y\\.y)?$")
  stats <- seasons_data %>% select(matches(pattern)) %>% as.matrix()
  weighted_mean <- rowSums(weights * stats, na.rm = TRUE) / rowSums(weights, na.rm = TRUE)
  return(weighted_mean)
}

all_playoffs$W_PCT <- weighted_avg(all_playoffs, "W_PCT")
all_playoffs$MIN <- weighted_avg(all_playoffs, "MIN")
all_playoffs$FGM <- weighted_avg(all_playoffs, "FGM")
all_playoffs$FGA <- weighted_avg(all_playoffs, "FGA")
all_playoffs$FG_PCT <- weighted_avg(all_playoffs, "FG_PCT")
all_playoffs$FG3M <- weighted_avg(all_playoffs, "FG3M")
all_playoffs$FG3A <- weighted_avg(all_playoffs, "FG3A")
all_playoffs$FG3_PCT <- weighted_avg(all_playoffs, "FG3_PCT")
all_playoffs$FTM <- weighted_avg(all_playoffs, "FTM")
all_playoffs$FTA <- weighted_avg(all_playoffs, "FTA")
all_playoffs$FT_PCT <- weighted_avg(all_playoffs, "FT_PCT")
all_playoffs$OREB <- weighted_avg(all_playoffs, "OREB")
all_playoffs$DREB <- weighted_avg(all_playoffs, "DREB")
all_playoffs$REB <- weighted_avg(all_playoffs, "REB")
all_playoffs$AST <- weighted_avg(all_playoffs, "AST")
all_playoffs$TOV <- weighted_avg(all_playoffs, "TOV")
all_playoffs$STL <- weighted_avg(all_playoffs, "STL")
all_playoffs$BLK <- weighted_avg(all_playoffs, "BLK")
all_playoffs$PF <- weighted_avg(all_playoffs, "PF")
all_playoffs$PFD <- weighted_avg(all_playoffs, "PFD")
all_playoffs$PTS <- weighted_avg(all_playoffs, "PTS")
all_playoffs$PLUS_MINUS <- weighted_avg(all_playoffs, "PLUS_MINUS")

all_playoffs <- all_playoffs %>%
  select('PLAYER_ID', 'PLAYER_NAME', 'TEAM_ID', 'TEAM_ABBREVIATION', 'AGE', 'total_GP', 'total_W', 'total_L', "W_PCT", "MIN", "FGM", "FGA", "FG_PCT", "FG3M", "FG3A", "FG3_PCT", "FTM", "FTA", "FT_PCT", "OREB", "DREB", "REB", "AST", "TOV", "STL", "BLK", "PF", "PFD", "PTS", "PLUS_MINUS")

```


```{r}
all_playoffs_clutch <- mutate(all_playoffs_clutch, clutch_PER = (85.910 * FGM + 53.897 * STL + 51.757 * FG3M + 46.845 * FTM + 39.190 * BLK + 39.190 * OREB + 34.677 * AST + 14.707 * DREB - 17.174 * PF - 20.091 * (FTA-FTM) - 39.190 * (FGA -FGM) - 53.897 * TOV)/MIN)

all_playoffs <- mutate(all_playoffs, playoff_PER = (85.910 * FGM + 53.897 * STL + 51.757 * FG3M + 46.845 * FTM + 39.190 * BLK + 39.190 * OREB + 34.677 * AST + 14.707 * DREB - 17.174 * PF - 20.091 * (FTA-FTM) - 39.190 * (FGA -FGM) - 53.897 * TOV)/MIN)

all_playoffs <- all_playoffs %>%
  filter(!is.na(playoff_PER))

playoff_elevation_df <- inner_join(updated_df, all_playoffs, by = "PLAYER_ID")

playoff_elevation_df <- playoff_elevation_df %>%
    mutate(playoff_elevation = playoff_PER - PER.x)

playoffs_df <- inner_join(all_playoffs_clutch, all_playoffs, by = "PLAYER_ID")

playoffs_df <- playoffs_df[playoffs_df$total_GP.x * playoffs_df$MIN.x >= 20, ]

cor_coef = cor(playoff_elevation_df$clutch_elevation, playoff_elevation_df$playoff_elevation)

ggplot(playoff_elevation_df , aes(x = clutch_elevation, y = playoff_elevation)) +
  geom_point() +  # Add points for the scatter plot
  geom_smooth(method = "lm", se = FALSE, color = "red") +  # Add a line of best fit
  labs(x = "Regular Season Clutch Elevation", y = "Playoff Elevation", title = "Clutch Elevation vs. Playoff Elevation") + 
  theme(plot.title = element_text(hjust = 0.5)) 

mean_value = mean(playoff_elevation_df$playoff_elevation)
sd_value = sd(playoff_elevation_df$playoff_elevation)

ggplot(playoff_elevation_df, aes(x = playoff_elevation)) + 
  geom_histogram(binwidth = 1, fill = "blue", color = "black") + # Adjust binwidth as needed
  ggtitle("Histogram of Playoff Elevation") + 
  xlab("Playoff Elevation") + 
  ylab("Frequency") +
  theme(plot.title = element_text(hjust = 0.5)) +
  annotate("text", x = Inf, y = Inf, label = paste("Mean =", round(mean_value, 2)), hjust = 1.1,vjust = 3, color="black", size = 4.5) +
  annotate("text", x = Inf, y = Inf, label = paste("SD =", round(sd_value, 2)), hjust = 1.1,vjust = 1.5, color="black", size = 4.5) +
  theme(plot.margin = margin(5.5, 40, 5.5, 5.5))
```


```{r}

playoffs_and_regular_df <- inner_join(updated_df, playoffs_df, by = "PLAYER_ID")
playoffs_and_regular_df <- playoffs_and_regular_df %>% 
  mutate(playoff_elevation = PER.y - PER.x) %>%
  mutate(clutch_playoff_elevation = clutch_PER.y - PER.x)

playoffs_and_regular_df %>%
  select(PLAYER_NAME.x.x, playoff_elevation) %>%
  arrange(desc(playoff_elevation))

playoffs_and_regular_df %>%
  select(PLAYER_NAME.x.x, clutch_playoff_elevation) %>%
  arrange(desc(clutch_playoff_elevation))

```







